---
title: "Landsim analysis for the Pereira et al. (2025) Adv Eco Res"
output: html_notebook
---
Remove old PopIn

```{r}
remove.packages("PopIn")
#.libPaths()
```


Install and load the library PopIn.

```{r}
#C++ code in PopIn was pre c++11 and followed the c++98 standard
#Have now updated it to compile with features that are deprecated in clang

remotes::install_github("anaceiahasse/PopIn", force=TRUE)
#remotes::install_local("/Users/henrique/Documents/Programming/labprojects/PopIn", force=TRUE,
#                       INSTALL_opts = c("--no-multiarch", "--with-keep.source","--no-byte-compile"))
remotes::install_local("/Users/henrique/Documents/Programming/labprojects/PopIn", force=TRUE)
```
Load the library PopIn and the fractal landscape generating function..

```{r}
library(PopIn)
source("FractalLandscape.R")

```
Now try an example. 

```{r}
indmodel()
```

Small home-range size

```{r}
land2 = generate_fractal_landscape(0.2, .01,20)
image(land2)
indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 1, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode = 0,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "small"))
```

Large home-range size fragmented.

```{r}
landfrag = generate_fractal_landscape(0.2, .01,160)
image(landfrag, asp=1)
indmodel(land_r=landfrag, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 64, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "large_fragmented"))
```
Large home-range size unfragmented (K=400).

```{r}
land2 = generate_fractal_landscape(0.2, .99,160)
image(land2, asp=1)
indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 64, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "large_unfrag"))
```
Large home-range size random (K=400).

```{r}
land2 = matrix(runif(160*160),160,160);
land2 = ConvertToLandscape(land2,.2)
image(land2, asp=1)
indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 64, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "large_random"))
```

Mechanistic home-range size unfragmented.

```{r}
land2 = generate_fractal_landscape(.2, .99,160)
image(land2, asp=1)
indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 0, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "mechanistic_unfrag"))
```
Mechanistic home-range size fragmented. 
```{r}
  indmodel(land_r=landfrag, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 0, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "mechanistic_fragmented"))
```
Mechanistic home-range size random 
```{r}
  land2 = matrix(runif(160*160),160,160);
  land2 = ConvertToLandscape(land2,.2)
  indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 0, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "FALSE"))
``` 

Mechanistic home-range size fragmented. 10% good habitat
```{r}
land2 = generate_fractal_landscape(0.1, .01,160)
image(land2, asp=1)
indmodel(land_r=land2, param = list(
  n_rep = 2, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 0, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "NO"))
```