---
title: "Landsim model analysis "
output: html_notebook
---

# About

-   Version: 1.1
-   Date: 2026-01-22
-   Author: Henrique M. Pereira
-   Paper: Pereira et al. (2025) Adv Eco Res

# Initialization

## Load needed packages

Remove old version of PopIn package (if needed).

```{r}
#remove.packages("PopIn")
```

Install the package PopIn.

```{r}
#C++ code in PopIn was pre c++11 and followed the c++98 standard
#Have now updated it to compile with features that are deprecated in clang

remotes::install_github("anaceiahasse/PopIn", ref="paper_resubmission")


#instal from local repo
#remotes::install_local("/Users/henrique/Documents/Programming/labprojects/PopIn", force=TRUE)

#install from local repo in debug mode
#remotes::install_local("/Users/henrique/Documents/Programming/labprojects/PopIn", #                       force=TRUE,
#                       INSTALL_opts = c("--no-multiarch", 
#                       "--with-keep.source","--no-byte-compile"))

packageVersion("PopIn")
```

Load the package PopIn and the fractal landscape generating function..

```{r}
library(PopIn)
source("FractalLandscape.R")

```

Now try an example.

```{r}
indmodel()
```

## Function to import model output

```{r}
# functions to read the model outputs that are based on Mathematica code (based on ChatGPT)
# originally the model output was Mathematica code

# -------------------------------
# Import Mathematica assignments to R
# -------------------------------

import_mathematica <- function(file) {
  lines <- readLines(file)
  full_text <- paste(lines, collapse = "\n")
  assignments <- unlist(strsplit(full_text, ";"))
  assignments <- trimws(assignments)
  assignments <- assignments[nzchar(assignments)]
  
  result <- list()
  
  parse_value <- function(value_str) {

  value_str <- trimws(value_str)

  # ---- CASE 1: rule-based assignment (simoptions) ----
  if (grepl("->", value_str)) {

    # remove ONLY the outer braces
    value_str <- sub("^\\{", "", value_str)
    value_str <- sub("\\}$", "", value_str)

    rules <- strsplit(value_str, ",")[[1]]
    lst <- list()

    for (r in rules) {
      kv <- strsplit(trimws(r), "->")[[1]]
      key <- trimws(kv[1])
      val <- trimws(kv[2])

      # numeric if possible, otherwise keep as character
      num <- suppressWarnings(as.numeric(val))
      lst[[key]] <- if (!is.na(num)) num else val
    }

    return(lst)
  }

  # ---- CASE 2: everything else (UNCHANGED logic) ----
  value_str <- gsub("\\{", "list(", value_str)
  value_str <- gsub("\\}", ")", value_str)
  value_str <- gsub("\\bTrue\\b", "TRUE", value_str)
  value_str <- gsub("\\bFalse\\b", "FALSE", value_str)

  obj <- tryCatch(eval(parse(text = value_str)), error = function(e) NULL)

  list_to_array <- function(x) {
    if (!is.list(x)) return(x)
    if (all(sapply(x, function(y) !is.list(y)))) {
      return(unlist(x))
    } else {
      x_conv <- lapply(x, list_to_array)
      if (all(sapply(x_conv, is.atomic)) &&
          length(unique(sapply(x_conv, length))) == 1) {
        return(do.call(rbind, x_conv))
      }
      return(x_conv)
    }
  }

  list_to_array(obj)
}


  for (assign in assignments) {
    if (!grepl("=", assign)) next
    
    # Detect sublist assignment: name[[i]] = value
    sublist_match <- regmatches(assign, regexec("^([a-zA-Z0-9_]+)\\[\\[([0-9]+)\\]\\]\\s*=\\s*(.*)$", assign))[[1]]
    
    if (length(sublist_match) > 0) {
      name <- sublist_match[2]
      index <- as.integer(sublist_match[3])
      value_str <- sublist_match[4]
      value <- parse_value(value_str)
      if (is.null(result[[name]])) result[[name]] <- vector("list", max(index,1))
      length(result[[name]]) <- max(length(result[[name]]), index)
      result[[name]][[index]] <- value
    } else {
      # Normal assignment
      parts <- strsplit(assign, "=")[[1]]
      name <- trimws(parts[1])
      value <- paste(parts[-1], collapse = "=")
      value <- trimws(value)
      result[[name]] <- parse_value(value)
    }
  }
  
  # Convert popsize (or any similar list of scalars) to vector
  if (!is.null(result$popsize)) {
    if (all(sapply(result$popsize, function(x) length(x) == 1))) {
      result$popsize <- unlist(result$popsize)
    }
  }
  
  return(result)
}


```

## Functions to plot home-ranges on landscape

```{r}
## Based on chagpt conversion from minpolygonplot in homerange.m

MinPolygon <- function(m, perc = 1, method = MinPolygonPerc) {
  method(m, perc)
}

MinPolygonPerc <- function(m, perc = 1) {

  n <- nrow(m)
  if (n == 0) return(m)

  ## DiscretizeCoords: grid cell centers (rounded)
  m0 <- cbind(round(m[,1]), round(m[,2]))

  ## Tally frequencies
  key <- paste(m0[,1], m0[,2], sep = "_")
  freq <- table(key)
  freq_df <- data.frame(
    cell = names(freq),
    freq = as.integer(freq),
    stringsAsFactors = FALSE
  )

  ## Recover coordinates
  coords <- do.call(rbind, strsplit(freq_df$cell, "_"))
  coords <- cbind(as.numeric(coords[,1]), as.numeric(coords[,2]))

  freq_df$x <- coords[,1]
  freq_df$y <- coords[,2]

  ## Sort ascending by frequency
  freq_df <- freq_df[order(freq_df$freq), ]

  ## CDF
  csums <- cumsum(freq_df$freq) / n

  idx <- which(csums <= 1 - perc)

  if (length(idx) == 0) {
    lpoints <- as.matrix(freq_df[, c("x","y")])
  } else {
    cp1 <- max(idx)
    freqcp <- freq_df$freq[cp1]

    tied <- which(freq_df$freq == freqcp)
    cp0 <- min(tied)
    cp2 <- max(tied)

    lA <- as.matrix(freq_df[tied, c("x","y")])
    lB <- as.matrix(freq_df[-seq_len(cp2), c("x","y")])

    ## Minimal area selection: approximate by including all
    lpoints <- rbind(lA, lB)
  }

  ## Final minimum polygon = convex hull
  if (nrow(lpoints) < 3) return(lpoints)

  hull <- chull(lpoints)
  lpoints[hull, , drop = FALSE]
}


MinPolygonPlot <- function(
  m,
  perc = 1,
  name = "",
  plot_style = "black",
  show_title = FALSE
) {

  if (length(m) == 0 || nrow(m) == 0) {
    return(list(poly = NULL, label = NULL, col = plot_style))
  }

  poly <- MinPolygon(m, perc)

  ## Close polygon
  poly <- rbind(poly, poly[1, ])

  label <- ""
  if (show_title) {
    label <- name
  }

  list(
    poly = poly,
    label = label,
    col = plot_style
  )
}


## Based on chatgpt conversion from Mathmetica LandscapeAnalysis,nb
PlotLandscape <- function(land, hrmapIn, randomcol=NULL) {

  if  (is.null(randomcol))
    randomcol = rainbow(length(hrmapIn))
  nr <- nrow(land)
  nc <- ncol(land)

  #oldpar <- par(no.readonly = TRUE)
  #on.exit(par(oldpar))

  par(pty = "s")   # force square plotting region

  plot(
    NA,
    xlim = c(1, nr),
    ylim = c(1, nc),
    asp = 1,
    xlab = "",
    ylab = "",
    frame.plot = TRUE,
    axes = FALSE
  )

  image(
    x = seq_len(nr),
    y = seq_len(nc),
    z = t(land),
    col = gray.colors(100, start = 1, end = 0),
    add = TRUE
  )

  box()

  for (i in seq_along(hrmapIn)) {

    mp <- MinPolygonPlot(
      m = hrmapIn[[i]],
      plot_style = randomcol[i]
    )

    if (!is.null(mp$poly)) {
      lines(
        mp$poly[,2],
        mp$poly[,1],
        col = mp$col,
        lwd = 2
      )
    }
  }
}



```

# Simulations

## Small home-range

```{r}
land2 = generate_fractal_landscape(0.2, .01,20)
image(land2)
indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 1, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode = 0,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "small"))
```

## Large home-range fragmented.

```{r}
landfrag = generate_fractal_landscape(0.2, .01,160)
image(landfrag, asp=1)
indmodel(land_r=landfrag, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 64, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "large_fragmented"))
```

## Large home-range unfragmented 

Note that K=400.

```{r}
land2 = generate_fractal_landscape(0.2, .99,160)
image(land2, asp=1)
indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 64, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "large_unfrag"))
```

## Large home-range random

```{r}
land2 = matrix(runif(160*160),160,160);
land2 = ConvertToLandscape(land2,.2)
image(land2, asp=1)
indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 64, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "large_random"))
```

## Mechanistic home-range unfragmented.

```{r}
land2 = generate_fractal_landscape(.2, .99,160)
image(land2, asp=1)
indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 0, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "mechanistic_unfrag"))
```

Read the output
```{r}
out = import_mathematica("Res_indmodel/mechanistic_unfrag_1")
out$simoptions
dim(out$landscape)
out$popsize
head(out$ageshist)
out$hrmaphist[[1]][[2]]
sapply(out$hrmaphist[[1]],dim)[1,]
sapply(out$hrmaphist[[31]],dim)[1,]
```
Plot population size and home-ranges
```{r}
plot(out$popsize, col = gray(0.5), pch = 16, xlab="Time", ylab="Population size")
```

```{r, fig.width=12, fig.height=4}
par(mfrow = c(1, 3), mar = c(2, 2, 2, 1))
PlotLandscape(out$landscape,out$hrmaphist[[1]])
PlotLandscape(out$landscape,out$hrmaphist[[16]])
PlotLandscape(out$landscape,out$hrmaphist[[31]])
```

## Mechanistic home-range size fragmented.

```{r}
  indmodel(land_r=landfrag, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 0, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "TRUE", 
  dir_name = "Res_indmodel", file_name = "mechanistic_fragmented"))
```

```{r}
out2 = import_mathematica("Res_indmodel/mechanistic_fragmented_27")
plot(out2$popsize, col = gray(0.5), pch = 16, xlab="Time", ylab="Population size",type="l")
PlotLandscape( out2$landscape,out2$hrmaphist[[1]])
```

## Mechanistic home-range size random

```{r}
  land2 = matrix(runif(160*160),160,160);
  land2 = ConvertToLandscape(land2,.2)
  indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 0, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "FALSE"))
```

## Mechanistic home-range size fragmented. 10% good habitat

```{r}
land2 = generate_fractal_landscape(0.1, .01,160)
image(land2, asp=1)
indmodel(land_r=land2, param = list(
  n_rep = 40, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
  hr_size = 0, birth_rate = 2.0, breeding_age = 1, survival = 0.4, 
  distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0 ,
  sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
  plotYN = "TRUE", writeYN = "NO"))
```

# Figures
## Appendix C
Figure C.1
```{r}
pdf("figures/fig_C1.pdf", width = 12, height = 8)
par(mfrow = c(2, 3), mar = c(2, 2, 2, 1))
PlotLandscape(out$landscape,out$hrmaphist[[1]])
PlotLandscape(out$landscape,out$hrmaphist[[16]])
PlotLandscape(out$landscape,out$hrmaphist[[31]])
PlotLandscape(out2$landscape,out2$hrmaphist[[1]])
PlotLandscape(out2$landscape,out$hrmaphist[[16]])
PlotLandscape(out2$landscape,out$hrmaphist[[31]])
dev.off()
```
Figure C.2
```{r}
pdf("figures/fig_C2.pdf", width = 8, height = 6)
plot(out$popsize, type="l", ylim=c(0,160),
     xlab="Time", ylab="Population size")
lines(out2$popsize,lty = 2)
dev.off()
```

Figure C.3
```{r}
library(future)
library(future.apply)
parallel::detectCores()
plan(multisession, workers = 10)
```

```{r}
nsteps=60
nrep=40
goodhab=seq(.999,0.01,-.05)
goodhab=c(goodhab,0.001)
popmatfrag=array(dim=c(length(goodhab),nrep,nsteps+1))
popmatunfrag=array(dim=c(length(goodhab),nrep,nsteps+1))
popmatrand=array(dim=c(length(goodhab),nrep,nsteps+1))
param_model=list(
      n_rep = 1, hab_aff = c(1.0, 0.1), n_steps = 60, init_population = 10,
      hr_size = 0, birth_rate = 2.0, breeding_age = 1, survival = 0.4,
      distance_weight = 0.001, dispersal_distance = 0, dispersal_mode =0,
      sink_avoidance = 0, neigh_avoidance = 0, sink_mortality = 0, 
      plotYN = "FALSE", writeYN = "FALSE")

run_one_i <- function(i) {
 capture.output({
  frag   <- matrix(NA, length(goodhab), nsteps + 1)
  unfrag <- matrix(NA, length(goodhab), nsteps + 1)
  rand   <- matrix(NA, length(goodhab), nsteps + 1)
  for (p in 1:length(goodhab))
    {
    land = generate_fractal_landscape(goodhab[p], .01,160)
    out=indmodel(land_r=land, param = param_model)
    frag[p, ] <- out$popsize_av_time_step
    
    land = generate_fractal_landscape(goodhab[p], .99,160)
    out=indmodel(land_r=land, param = param_model)
    unfrag[p, ] <- out$popsize_av_time_step
    
    land = matrix(runif(160*160),160,160);
    land = ConvertToLandscape(land,goodhab[p])
    out=indmodel(land_r=land, param = param_model)
    rand[p, ] <- out$popsize_av_time_step
  }
 }, file = NULL)
list(
    i = i,
    frag   = frag,
    unfrag = unfrag,
    rand   = rand)  
}

results <- future_lapply(
  seq_len(nrep),
  run_one_i,
  future.seed = TRUE
)

for (res in results) {
  popmatfrag[,res$i,]   <- res$frag
  popmatunfrag[,res$i,] <- res$unfrag
  popmatrand[,res$i,]   <- res$rand
}

save(popmatfrag,popmatunfrag,popmatrand,file="Res_indmodel/figurec3.RData")
```

Plots
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

time_window <- 51:61
P <- dim(popmatfrag)[1]
nrep <- dim(popmatfrag)[2]

# Helper function to summarize an array
summarize_array <- function(arr, label) {
  df <- as.data.frame.table(arr)  # gives Var1, Var2, Var3, Freq
  colnames(df) <- c("p", "i", "t", "popsize")
  df$p <- as.numeric(df$p)
  df$i <- as.numeric(df$i)
  df$t <- as.numeric(df$t)
  
  df %>%
    filter(t %in% time_window) %>%
    group_by(p, i) %>%
    summarize(mean_time = mean(popsize)/4, .groups = "drop") %>%
    group_by(p) %>%
    summarize(
      mean_pop = mean(mean_time),
      q25 = quantile(mean_time, 0.25),
      q75 = quantile(mean_time, 0.75),
      .groups = "drop"
    ) %>%
    mutate(type = label)
}

# Summarize all three
df_frag   <- summarize_array(popmatfrag,   "frag")
df_unfrag <- summarize_array(popmatunfrag, "unfrag")
df_rand   <- summarize_array(popmatrand,   "rand")

# Combine into one dataframe
df_plot <- bind_rows(df_frag, df_unfrag, df_rand)

# Replace the p indices with actual habitat values
df_plot <- df_plot %>%
  mutate(p = ((1- goodhab[p])*100))

myplot=
  ggplot(df_plot, aes(x = p, y = mean_pop, linetype = type, group = type)) +
  geom_line(size = 1) +
  geom_point(aes(shape = type), size = 2) +
  geom_errorbar(aes(ymin = q25, ymax = q75), width = 0.2) +
  scale_linetype_manual(values = c("dotted","dashed","solid"), name=NULL) +
  scale_shape_manual(values = c(16, 17, 15),name = NULL) +  #symbols
  labs(
    x = "Habitat loss (%)",
    y = "Realized carrying capacity (%)",
    linetype = "Landscape type",
    shape = "Landscape type"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top")
myplot
# Save as PDF
ggsave("figures/fig_C3.pdf", plot = myplot, width = 7, height = 5)  
```
```{r}
popmatrand[20,,61]
popmatunfrag[20,,61]
popmatfrag[20,,61]
```




